### 2022-03-08 paper/localpsf.tex
paper edits

Our method is based on approximation \eqref{eq:translate3}, which is an improvement upon \eqref{eq:translate1} and \eqref{eq:translate2} that generalizes both. We call \eqref{eq:translate3} ``local mean-displacement invariance,'' and illustrate \eqref{eq:translate3} in Figure \ref{fig:mean_displacement_invariance}. On one hand, if \eqref{eq:translate1} holds, then $\mu(x) \approx \mu(x')$, and so \eqref{eq:translate3} holds. On the other hand, translating a distribution translates its mean, so if \eqref{eq:translate2} holds, then $\mu(x')-\mu(x) \approx x' - x$, so again \eqref{eq:translate3} holds. But approximation \eqref{eq:translate3} can hold in situations where neither \eqref{eq:translate1} nor \eqref{eq:translate2} holds. For example, because the expected value is a linear operator, \eqref{eq:translate3} will hold when an $\Aop$ is locally translation-invariant with respect to a rotated frame of reference, while \eqref{eq:translate2} will not hold in this case. Additionally, \eqref{eq:translate3} generalizes to operators $\Aop:L^2(\Omega_1) \rightarrow L^2(\Omega_2)'$ that map between function spaces on different domains $\Omega_1$ and $\Omega_2$, and even operators that map between domains with different spatial dimensions. In contrast, \eqref{eq:translate2} does not naturally generalize to operators that map between function spaces on different domains, because the formula $y-x+x'$ requires vectors in $\Omega_2$ and $\Omega_1$ to be added together.

@@

%In Section \ref{sec:intro} we noted that computing $\Aker\left(y,x\right)$ is computationally costly. 
%This section describes how we compute approximate kernel entries
%\begin{equation*}
%\widetilde{\Phi}(y,x) \approx \Phi(y,x)
%\end{equation*}
%for arbitrary pairs of points $(y,x) \in \Omega \times \Omega$. 
Given $(y,x)\in \Omega \times \Omega$, let
\begin{equation}
\label{eq:fxyxp}
f_i := \frac{V(x)}{V(x_i)}\phi_{x_i}\left(y - \mu(x) + \mu(x_i)\right), \qquad i=1,\dots,k,
\end{equation}
%for arbitrary pairs of points $(y,x) \subset \Omega \times \Omega$.
% while Section \ref{sec:approximate_kernel_entries} we showed how we can cheaply evaluate approximate kernel entries at arbitrary pairs of points $(y,x) \in \Omega \times \Omega$.
where $\{x_i\}_{i=1}^k$ are the $k$ nearest sample points to $x$, excluding sample points for which $\phi_{x_i}\left(y - \mu(x) + \mu(x_i)\right)$ is undefined. That is, excluding sample points such that $y-\mu(x)+\mu(x_i) \notin \Omega$. The number $k$ is a small user defined parameter; in numerical experiments we use $k=10$. We find the $k$ nearest sample points by querying a precomputed KD-tree \cite{kdtree} of all sample points. We check whether $y - \mu(x) + \mu(x_i) \in \Omega$ by querying a precomputed axis aligned bounding box tree (AABB tree) \cite{aabbtree} of the mesh cells used to discretize the problem.

The idea here is that $\Phi(y,x) \approx f_i$ when $x_i$ is near $x$, so we approximate $\Phi(y,x)$ by interpolating the (point,value) pairs
\begin{equation}
\left(x_i, f_i\right), \qquad i=1,\dots,k,
\end{equation}
at the point $x$. To see that $\Phi(y,x) \approx f_i$, recall from Section \ref{eq:local_mean_displacement_invariance} that $\Phi(y,x) = \phi_x(y) \approx \phi_{x_i}\left(y - \mu(x) + \mu(x_i)\right)$ if $x_i$ is near $x$. Inclusion of the extra factor $\frac{V(x)}{V(x_i)}$ in \eqref{eq:fxyxp} is a form of preconditioning that makes this approximation more robust in the case where $V(x)$ varies widely. 
%Note that we are not in danger of dividing by zero, because our ellipsoid packing procedure from Section \ref{sec:sample_point_selection} excludes $x_i$ from consideration as a sample point if $V(x_i)$ is smaller than a predetermined threshold.

%\subsubsection{Radial basis function interpolation}

We use the following radial basis function interpolation scheme:

@@

We use polyharmonic spline radial basis functions \cite{polyspline}, which are defined as follows:

@@

for $r>0$, and $b_i(0):=0$. Polyharmonic spline radial basis function interpolation yields the smoothest interpolant of the given points and values, in a certain least-squares sense. For more details on radial basis function interpolation and polyharmonic splines, see \cite{rbf}.

>>>>

Our method is based on approximation \eqref{eq:translate3}, which is an improvement upon \eqref{eq:translate1} and \eqref{eq:translate2} that generalizes both. We call \eqref{eq:translate3} ``local mean-displacement invariance,'' and illustrate \eqref{eq:translate3} in Figure \ref{fig:mean_displacement_invariance}. On one hand, if \eqref{eq:translate1} holds, then $\mu(x) \approx \mu(x')$, and so \eqref{eq:translate3} holds. On the other hand, translating a distribution translates its mean, so if \eqref{eq:translate2} holds, then $\mu(x')-\mu(x) \approx x' - x$, so again \eqref{eq:translate3} holds. But approximation \eqref{eq:translate3} can hold in situations where neither \eqref{eq:translate1} nor \eqref{eq:translate2} holds. For example, because the expected value commutes with affine transformations, \eqref{eq:translate3} will hold when an $\Aop$ is locally translation-invariant with respect to a rotated frame of reference, while \eqref{eq:translate2} will not hold in this case. Additionally, \eqref{eq:translate3} generalizes to operators $\Aop:L^2(\Omega_1) \rightarrow L^2(\Omega_2)'$ that map between function spaces on different domains $\Omega_1$ and $\Omega_2$, and even operators that map between domains with different spatial dimensions. In contrast, \eqref{eq:translate2} does not naturally generalize to operators that map between function spaces on different domains, because the formula $y-x+x'$ requires vectors in $\Omega_2$ and $\Omega_1$ to be added together.

@@

%Recall from Section \ref{eq:local_mean_displacement_invariance} that if $x'$ is near $x$, then
%\begin{equation*}
%\Phi(y,x) = \phi_x(y) \approx \phi_{x_i}\left(y - \mu(x) + \mu(x')\right).
%\end{equation*}

%In Section \ref{sec:intro} we noted that computing $\Aker\left(y,x\right)$ is computationally costly. 
%This section describes how we compute approximate kernel entries
%\begin{equation*}
%\widetilde{\Phi}(y,x) \approx \Phi(y,x)
%\end{equation*}
%for arbitrary pairs of points $(y,x) \in \Omega \times \Omega$. 
Given $(y,x)\in \Omega \times \Omega$, let
\begin{equation}
\label{eq:fxyxp}
f_i := \frac{V(x)}{V(x_i)}\phi_{x_i}\left(y - \mu(x) + \mu(x_i)\right), \qquad i=1,\dots,k,
\end{equation}
%for arbitrary pairs of points $(y,x) \subset \Omega \times \Omega$.
% while Section \ref{sec:approximate_kernel_entries} we showed how we can cheaply evaluate approximate kernel entries at arbitrary pairs of points $(y,x) \in \Omega \times \Omega$.
where $\{x_i\}_{i=1}^k$ are the $k$ nearest sample points to $x$, excluding sample points for which $\phi_{x_i}\left(y - \mu(x) + \mu(x_i)\right)$ is undefined. We approximate $\Phi(y,x)$ by interpolating the (point,value) pairs
\begin{equation}
\left(x_i, f_i\right), \qquad i=1,\dots,k,
\end{equation}
at the point $x$. The idea here is that $\Phi(y,x) = \phi_x(y) \approx \phi_{x_i}\left(y - \mu(x) + \mu(x_i)\right)$ (see Section \ref{eq:local_mean_displacement_invariance}), so we approximate $\Phi(y,x)$ by interpolating the values $\phi_{x_i}\left(y - \mu(x) + \mu(x_i)\right)$. Inclusion of the factor $\frac{V(x)}{V(x_i)}$ in \eqref{eq:fxyxp} is a form of preconditioning that makes this approximation more robust in the case where $V(x)$ varies widely.

To find the $k$ nearest sample points to $x$, we query a precomputed KD-tree \cite{kdtree} of all sample points. The number $k$ is a small user defined parameter; in numerical experiments we use $k=10$. The function $\phi_{x_i}\left(y - \mu(x) + \mu(x_i)\right)$ is defined if and only if $y-\mu(x)+\mu(x_i) \in \Omega$. We check whether $y - \mu(x) + \mu(x_i) \in \Omega$ by querying a precomputed axis aligned bounding box tree (AABB tree) \cite{aabbtree} of the mesh cells used to discretize the problem. 
%Since $\Phi(y,x) \approx f_i$ when $x_i$ is near $x$, we approximate $\Phi(y,x)$ by interpolating the (point,value) pairs
%\begin{equation}
%\left(x_i, f_i\right), \qquad i=1,\dots,k,
%\end{equation}
%at the point $x$. To see that $\Phi(y,x) \approx f_i$, recall from Section \ref{eq:local_mean_displacement_invariance} that $\Phi(y,x) = \phi_x(y) \approx \phi_{x_i}\left(y - \mu(x) + \mu(x_i)\right)$ if $x_i$ is near $x$. Inclusion of the extra factor $\frac{V(x)}{V(x_i)}$ in \eqref{eq:fxyxp} is a form of preconditioning that makes this approximation more robust in the case where $V(x)$ varies widely. 
%Note that we are not in danger of dividing by zero, because our ellipsoid packing procedure from Section \ref{sec:sample_point_selection} excludes $x_i$ from consideration as a sample point if $V(x_i)$ is smaller than a predetermined threshold.
%\subsubsection{Radial basis function interpolation}

For interpolation, we use the following radial basis function scheme:

@@

We use polyharmonic spline radial basis functions \cite{polyspline}:

@@

for $r>0$, and $b_i(0):=0$. Polyharmonic spline radial basis function interpolation yields the smoothest interpolant of the given points and values, in a least-squares sense. For more details on radial basis function interpolation and polyharmonic splines, see \cite{rbf}.
